---
title: "The Effect of Weather Events on Population Health and Economic Consequences"
author: "YK Zhao"
date: "Saturday, June 20, 2015"
output:
  pdf_document: default
  html_document:
    fig_caption: yes
    theme: spacelab
---
##0 Title: The Effect of Weather Events on Population Health and Economic Consequences

##1 Synopsis
The basic goal of this project is to explore the NOAA Storm Database and answer some basic questions about severe weather events. Questions are (1) Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health? and (2) Across the United States, which types of events have the greatest economic consequences?

Results from data and figures show that "Tornado" leads to highest fatality counts and "Tornado" leads to highest injury counts. "Tornado" causes most severe property damage and "Drought" causes most severe crop damage.

##2 Set Environment
```{r setEnvironment}
##Set environment
##setwd("C:/Users/User/Desktop/Reproducible research/Peer assessment 2/") 
library(knitr)
library(ggplot2)
library(plyr)
opts_chunk$set(echo=TRUE)
```

##3 Data Processing
##First you should download data, unzip data and load data into R
```{r loadDta}
##Download zip file
fileUrl <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(fileUrl, destfile = "./repdata-data-StormData.csv.bz2")
timeDownload <- date()

##Load csv file after unzipping
##unzip("./repdata-data-StormData.zip")
storm <- read.csv("./repdata-data-StormData.csv", header=TRUE)
##head(storm, 10)
```
Time for downloading zip file is `r timeDownload`.

##Then you should do some data processing
##Since the dataset is too large, it is necessary to select only those we need for this project
```{r select}
select <- c("EVTYPE","FATALITIES","INJURIES","PROPDMG", "PROPDMGEXP", "CROPDMG", "CROPDMGEXP") ##Variables we need
storm <- storm[1:547363, select] ##Rows we need
head(storm, 10)
str(storm)
summary(storm)
```

##Population health
```{r PH}
##Get fatality and injury
fatality <- storm[1:547363, c("EVTYPE","FATALITIES")]
injury <- storm[1:547363, c("EVTYPE","INJURIES")]

##Aggregate sum them by event
fatality <- aggregate(as.numeric(as.character(fatality$FATALITIES)), by=list(fatality$EVTYPE), sum)
injury <- aggregate(as.numeric(as.character(injury$INJURIES)), by=list(injury$EVTYPE), sum)
names(fatality) <- c("event","fatalitycount")
names(injury) <- c("event","injurycount")
```

##Economic consequence
##In this dataset, economic consequence is measured by property damage and crop damage
```{r EC, cache=TRUE}
##Get property damage: Need transform with exponent unit
##K for thousands, M for millions, B for billions
propdmg <- storm[1:547363, c("EVTYPE","PROPDMG","PROPDMGEXP")]
propdmg$prop <- 0 ##Initialize
propdmg$PROPDMGEXP <- as.character(propdmg$PROPDMGEXP)
propdmg$PROPDMG <- as.numeric(as.character(propdmg$PROPDMG))

for (i in 1:nrow(propdmg)) { ##Transform with exponent unit
  if(toupper(propdmg[i,3]) == "K") {
    propdmg[i,4] <- propdmg[i,2]*10^3
  }else if(toupper(propdmg[i,3]) == "M") {
    propdmg[i,4] <- propdmg[i,2]*10^6
  }else if(toupper(propdmg[i,3]) == "B") {
    propdmg[i,4] <- propdmg[i,2]*10^9
  }else {
    propdmg[i,4] <- propdmg[i,2]
  }##End if
}##End for

##Get crop damage: Need transform with exponent unit
##K for thousands, M for millions, B for billions
cropdmg <- storm[1:547363, c("EVTYPE","CROPDMG","CROPDMGEXP")]
cropdmg$crop <- 0 ##Initialize
cropdmg$CROPDMGEXP <- as.character(cropdmg$CROPDMGEXP)
cropdmg$CROPDMG <- as.numeric(as.character(cropdmg$CROPDMG))

for (i in 1:nrow(cropdmg)) { ##Transform with exponent unit
  if(toupper(cropdmg[i,3]) == "K") {
    cropdmg[i,4] <- cropdmg[i,2]*10^3
  }else if(toupper(cropdmg[i,3]) == "M") {
    cropdmg[i,4] <- cropdmg[i,2]*10^6
  }else if(toupper(cropdmg[i,3]) == "B") {
    cropdmg[i,4] <- cropdmg[i,2]*10^9
  }else {
    cropdmg[i,4] <- cropdmg[i,2]
  }##End if
}##End for

##Aggregate sum them by event
propdmg <- aggregate(propdmg$prop, by=list(propdmg$EVTYPE), sum)
cropdmg <- aggregate(cropdmg$crop, by=list(cropdmg$EVTYPE), sum)
names(propdmg) <- c("event","propertydamage")
names(cropdmg) <- c("event","cropdamage")
```

##4 Results
##First get top 10 events for each measure (fatality, injury, property damage and crop damage)
```{r top10}
##Population health
fatality10 <- arrange(fatality, desc(fatalitycount))[1:10,]
injury10 <- arrange(injury, desc(injurycount))[1:10,]

##Economic consequences
propdmg10 <- arrange(propdmg, desc(propertydamage))[1:10,]
cropdmg10 <- arrange(cropdmg, desc(cropdamage))[1:10,]

##Question 1 Which event is most harmful with respect to population health?
fatality10[1,]
injury10[1,]

##Question 2 Which event is most harmful with respect to economic consequences?
propdmg10[1,]
cropdmg10[1,]

##Reshape and append datasets to append in terms of population health and economic consequences
fatality10$factor <- "fatality count"
injury10$factor <- "injury count"
names(fatality10) <- c("event","count","factor")
names(injury10) <- c("event","count","factor")
PH <- rbind(fatality10, injury10)

propdmg10$factor <- "property damage"
cropdmg10$factor <- "crop damage"
names(propdmg10) <- c("event","amount","factor")
names(cropdmg10) <- c("event","amount","factor")
EC <- rbind(propdmg10, cropdmg10)

##Plot figure for population health and economic consequences
p1 <- ggplot(data=PH, aes(x=event, y=count, group=factor(factor)))
p1 + geom_point(size=5, pch=21, fill="salmon", alpha=0.5) + labs(title="Top 10 Weather Events Harmful to Population Health", y="Counts", x="Events", fill="Measure Type")

p2 <- ggplot(data=EC, aes(x=event, y=amount, group=factor(factor)))
p2 + geom_point(size=5, pch=21, fill="salmon", alpha=0.5) + labs(title="Top 10 Weather Events Harmful to Economic Consequences", y="Counts", x="Events", fill="Measure Type")
```

It is obvious that in terms of population health, "Tornado" leads to highest fatality counts and "Tornado" leads to highest injury counts. In terms of economic consequences, "Tornado" causes most severe property damage and "Drought" causes most severe crop damage.
  
##5 Conclusions
Results from data and figures show that "Tornado" is most harmful weather events with respect to population health, and "Tornado" and "Drought" weather events have the greatest economic consequences.

